---
title: "Titanic"
author: "Mengyao Shi"
date: "March 22, 2017"
output: ioslides_presentation
---


```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(dplyr)
train<-read.csv('train.csv',stringsAsFactors = F)
test<-read.csv('test.csv',stringsAsFactors = F)
full<- bind_rows(train, test) 
```

```{r, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
library('ggthemes')
library(caret)
library(randomForest)
library(ElemStatLearn)
library('mice')
full$Title<-gsub('(.*, )|(\\..*)', '', full$Name)
other_title<-c('Capt','Col', 'Don', 'Dr','Jonkheer','Lady','Major','Rev','Sir','the Countess')
full$Title[full$Title == "Mlle"]<-'Miss'
full$Title[full$Title == 'Mme']<-'Mrs'
full$Title[full$Title == 'Ms']<-'Miss'
full$Title[full$Title %in% other_title]<-'Other Titles'
full$Surname <- sapply(full$Name,  function(x) strsplit(x, split = '[,.]')[[1]][1])
full$FSize<-full$SibSp+full$Parch+1
full$FullFamily<-paste(full$Surname, full$FSize, sep='_')
full$surviveRatio<-full$Survived



##missing value full[1044,] fare

full$Fare[1044] <- median(full[full$Pclass == '3' & full$Embarked == 'S', ]$Fare, na.rm = TRUE)
##missing value age
factor_vars <- c('PassengerId','Pclass','Sex','Embarked',
                 'Title','Surname','FullFamily','FSize')

full[factor_vars] <- lapply(full[factor_vars], function(x) as.factor(x))
set.seed(129)

miceOutput<-mice(full[!names(full) %in% c('PassengerId','Name','Ticket','Cabin','FullFamily','Surname','Survived','surviveRatio')],method='rf') %>%
  complete() 
full$Age <- miceOutput$Age

##finished
train<-full[1:891,]
test<-full[892:1309,]
comp_train<-train[complete.cases(train),]


split_By_Survive<-split(comp_train, comp_train$Survived)
Split_By_Sex<-split(comp_train, comp_train$Sex)
Split_By_Class<-split(comp_train,comp_train$Pclass)

factor_vars <- c('PassengerId','Pclass','Sex','Embarked',
                 'Title','Surname','FullFamily','FSize')

comp_train[factor_vars] <- lapply(comp_train[factor_vars], function(x) as.factor(x))

inTrain<-createDataPartition(y=comp_train$Survived,p=0.6,list=FALSE)
training<-comp_train[inTrain,]
validation<-comp_train[-inTrain,]

set.seed(876)
modelFit<-randomForest(factor(Survived)~Pclass+Sex+Age+SibSp+Parch+Fare+Embarked+Title+FSize,data=training)

```



```{r,echo=FALSE}
# Get importance
importance    <- importance(modelFit)
varImportance <- data.frame(Variables = row.names(importance), 
                            Importance = round(importance[ ,'MeanDecreaseGini'],2))

# Create a rank variable based on importance
rankImportance <- varImportance %>%
  mutate(Rank = paste0('#',dense_rank(desc(Importance))))

rankImportance<- rankImportance[with(rankImportance, order(-rankImportance$Importance)),]
# Use ggplot2 to visualize the relative importance of variables
```


## 891 people Survived and Dead statistics
### zero being dead and one means survived
```{r}
library(plotly)
plot_ly(comp_train, x=~Survived, type='histogram')
```

## First Class Passenger and Survival
```{r, echo=FALSE, message=FALSE}
library(plotly)
Class_3<-c("First Class", "Second Class", "Third Class")
plot_ly(Split_By_Class$`1`, x=~Survived, type='histogram')



```

## Second Class Passenger and Survival
```{r, echo=FALSE, message=FALSE}
library(plotly)
plot_ly(Split_By_Class$`2`, x=~Survived, type='histogram')
```

## Third Class Passenger and Survival
```{r, echo=FALSE, message=FALSE}
library(plotly)
plot_ly(Split_By_Class$`3`, x=~Survived, type='histogram')
```

## Ticket Fare
```{r}
library(plotly)
plot_ly(alpha=0.6) %>%
  add_histogram(x=split_By_Survive$`0`$Fare, type='histogram', name='Dead') %>%
  add_histogram(x=split_By_Survive$`1`$Fare, type='histogram', name='survived') %>%
  layout(barmode="stack", xaxis=list(title='Fare in USD'))

```

## Age and Survival
### Notice orange(survived) and blue(all) use a bit different binning 
```{r, message=FALSE}
library(plotly)

plot_ly(alpha = 0.6) %>%
  add_histogram(x=comp_train$Age, type="histogram", name='all') %>%
  add_histogram(x=split_By_Survive$`1`$Age, type="histogram", name='survived') %>%
  layout(barmode = "overlay", xaxis=list(title='Fare in USD'))

```

## Female and Survival
```{r, echo=FALSE, message=FALSE}
library(plotly)
Split_By_Sex<-split(comp_train, comp_train$Sex)
plot_ly(Split_By_Sex$female, x=~Survived, type='histogram')
```

## male Survival
```{r, echo=FALSE, message=FALSE}
library(plotly)
Split_By_Sex<-split(comp_train, comp_train$Sex)
plot_ly(Split_By_Sex$male, x=~Survived, type='histogram')
```



```{r}

modelFit<-randomForest(factor(Survived)~Pclass+Sex+Age+Fare+Title+FSize,data=training, ntree=40)
result_train<-predict(modelFit,training)
result_predict<-predict(modelFit,validation)
validation[,(ncol(validation)+1)]<-result_predict


```
## Relative importance of key features 
### Live or death? These are key factors that determin(this is not an interactive plot)
```{r}
# Use ggplot2 to visualize the relative importance of variables
ggplot(rankImportance, aes(x = reorder(Variables, Importance), 
    y = Importance, fill = Importance)) +
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
    hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_few()

```



```{r}
modelFit<-randomForest(factor(Survived)~Pclass+Sex+Age+Fare+Title+FSize,data=comp_train, ntree=40)
result_train<-predict(modelFit,comp_train)
prediction<-predict(modelFit,test)
# Save the solution to a dataframe with two columns: PassengerId and Survived (prediction)
solution <- data.frame(PassengerID = test$PassengerId, Survived = prediction)

# Write the solution to file
write.csv(solution, file = 'rf_Solution.csv', row.names = F)

```


## Random Forest prediction result(not an interactive plot)

```{r}
library(plotly)
plot(modelFit, ylim=c(0,0.36))
legend('topright', colnames(modelFit$err.rate), col=1:3, fill=1:3)
```

## Thank you
### Reference 
#### https://www.kaggle.com/c/titanic/data
#### https://www.kaggle.com/mrisdal/titanic/exploring-survival-on-the-titanic
